---
title: "Configuration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Configuration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

Before you can use egfiles to interact with Egnyte, you need to set up your credentials. This vignette walks through how to get an API key and configure it for use with egfiles. If you're looking to understand the different authentication methods available, check out `vignette("authorization")` after you've finished here.

## Getting an API Key

First things first - you need an API key from Egnyte. The process for obtaining one depends on whether you're an Egnyte administrator or a regular user.

### If You're an Admin

If you have admin access to your Egnyte account, you can generate an API key directly:
1. Log in to your Egnyte account
2. Navigate to **Settings** > **Security & Authentication** > **API Keys**
3. Click **Create New Key**
4. Give the key a descriptive name (e.g., "R egfiles access")
5. Copy the generated key - you won't be able to see it again

### If You're Not an Admin

If you don't have admin access, you'll need to request an API key from your Egnyte administrator. Send them a request explaining:
- What you need the API key for (R-based file access)
- What permissions you need (typically read/write access to specific folders)

Alternatively, you can use OAuth authentication, which doesn't require admin involvement. See `vignette("authorization")` for details on the OAuth options.

## Setting Up Your Credentials

Once you have your API key, you can configure egfiles using the `eg_auth()` function:

```{r}
library(egfiles)

eg_auth(
  domain = "your-company",
  api_key = "your-api-key-here"
)
```

The `domain` parameter is your Egnyte subdomain - the part that comes before `.egnyte.com` in your Egnyte URL. So if you access Egnyte at `https://acme-corp.egnyte.com`, your domain is `"acme-corp"`.

After calling `eg_auth()`, you're ready to use the file reading and writing functions.

## Using Environment Variables

Hardcoding credentials in your scripts is generally a bad idea. Anyone who sees your code also sees your API key. Instead, egfiles supports reading credentials from environment variables.

Set these environment variables on your system:
- `EGNYTE_DOMAIN`: Your Egnyte domain
- `EGNYTE_API_KEY`: Your API key

Then you can call `eg_auth()` without any arguments:

```{r}
eg_auth()
```

egfiles will automatically pick up the values from your environment.

### Setting Environment Variables

How you set environment variables depends on your operating system and workflow.

**In an .Renviron file (recommended for R users):**

Create or edit a file called `.Renviron` in your home directory:

```
EGNYTE_DOMAIN=your-company
EGNYTE_API_KEY=your-api-key-here
```

Restart R, and these variables will be available in every session.

**On macOS/Linux (terminal):**

Add to your `~/.bashrc`, `~/.zshrc`, or equivalent:

```bash
export EGNYTE_DOMAIN="your-company"
export EGNYTE_API_KEY="your-api-key-here"
```

**On Windows (Command Prompt):**

```cmd
setx EGNYTE_DOMAIN "your-company"
setx EGNYTE_API_KEY "your-api-key-here"
```

## Verifying Your Setup

After configuring your credentials, you can verify everything is working by trying to read a file you know exists:

```{r}
# Try reading a file from a known path
eg_read("/Shared/some-folder/test-file.txt", destfile = "test.txt")
```
If you get an error about authentication, double-check that:

1. Your API key is correct (no extra spaces or characters)
2. Your domain is correct (just the subdomain, not the full URL)
3. Your API key has the necessary permissions for the folder you're accessing

## A Note on Security

Your API key provides access to your Egnyte files. Treat it like a password:

- Never commit it to version control
- Don't share it in Slack, email, or other communication channels
- Use environment variables or a secrets manager rather than hardcoding
- If you suspect your key has been compromised, revoke it immediately and generate a new one

For team environments where you need more control over authentication and token management, consider using OAuth instead. See `vignette("authorization")` for details.

## Next Steps

Now that you have your credentials configured, you can:

- Learn about the different authentication methods in `vignette("authorization")`
- Start transferring files with `vignette("file-transfer")`
- Read and write data files with `vignette("reading-writing")`
